---
title: "Data Objects and Basic Visualization Activity"
output: github_document
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, size = "small", collapse=TRUE, comment=NA, warning = FALSE, message = FALSE)
```


```{r}
babynames <- read.csv("https://raw.githubusercontent.com/deepbas/statdatasets/main/baby-names-by-state.csv")
x <- c(3,6,9,5,10)
x.mat <- cbind(x, 2*x)
x.df <- data.frame(x=x,double.x=x*2)
my.list <- list(myVec=x, myDf=x.df, myString=c("hi","bye"))
```

### Question 1: data types

- What data type is `x`? What data type is `babynames$number`? 
- What data type is `c(x, babynames$year)`?
- What data type is `c(x,NA)`?  What data type is `c(x,"NA")`?

### Question 2: Subsetting and coercion

- How can we reverse the order of entries in `x`?
- What does `which(x < 5)` equal?
- What does `sum(c(TRUE,FALSE,TRUE,FALSE))` equal?
- What does `sum(x[c(TRUE,FALSE,TRUE,FALSE)])` equal?
- What does `sum(x < 5)` equal?
- What does `sum(x[x < 5])` equal?
- Why `dim(x.mat2[1:2,1])` return `NULL` while `dim(x.mat2[1:2,1:2])` returns a dimension?

### Question 3: Lists

- Using `my.list`, show three ways to write one command that gives the 3rd entry of variable `x` in data frame `myDf`
- What class of object does the command `my.list[3]` return?
- What class of object does the command `my.list[[3]]` return?
- What class of object does the command `unlist(my.list)` return? Why are all the entries `character`s?

### Question 4: Data Frames

Loading the `tidyverse` loads 8 packages, one of which is `ggplot2`. You can certainly load each package individually, but it has become common to simply load the `tidyverse`.

```{r}
#install.packages("tidyverse")
library(tidyverse)
babynames_MN_Amy <- babynames %>% filter(state=="MN",  name == "Amy")
```

- What is the total number of babies born in Minnesota with name `John`?

```{r}
babynames_MN_John <- babynames %>% filter(state=="MN",  name == "John")

```


- In what year were highest number of babies were named Alex in Minnesota?

### Question 5: Basic Plot using `ggplot2`.

```{r}
#install.packages("ggplot2")
library(ggplot2)
```

- What are the grammar of graphics needed to create a scatter-plot of the number of babies born in Minnesota named `John` vs year they were born.

- **data**: 
- **aesthetic mapping**: `y = \ldots   `, `x =  \ldots`
- **geometric object**: 

- Fill in the data and aesthetic mapping in the below code chunk. What is returned? What's missing?

```{r, eval = FALSE}
# Fill in the blanks
ggplot(data = babynames_MN_John , mapping = aes(x = year, y = number ))
```

- Add the appropriate geometric object to create the scatter plot. This is called adding a *layer* to a plot. Remember to always put the `+` at the end of a line, never at the start.


```{r}
ggplot(data = babynames_MN_John , mapping = aes(x = year, y = number )) +
  geom_point()

```

- Repeat the above steps with babies named `John` from Minnesota or Michigan.

```{r}
babynames_MN_MI_John <- babynames %>% filter(state=="MN" | state == "MI",  name == "John")
ggplot(data = babynames_MN_MI_John , mapping = aes(x = year, y = number )) +
  geom_point(aes(colour=state))
```

- Finally, add appropriate labels

```{r fig.cap= "A trend chart", tidy=FALSE, eval=FALSE}
ggplot(data = babynames_MN_MI_John , mapping = aes(x = year, y = number)) +
  geom_point(aes(colour=state))+
  xlab("Year") +
  ylab(bquote(Number~of~babies~named~.("John")))
```